Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-06-16 code/rmarkdown/article.Rmd
+ update article with population changes + change endnote_lib.bib to references.bib

---
title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: false
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
      after_body: postamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
---

@@

## Methods
### Study system
We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

@@

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation. We obtained data for 19 bioclimatic variables across the extent of the study area [approx. 1 km resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

@@

The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456]. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive. To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; XXXX] assuming various numbers of populations (K = `r min(structure.params.LST[[MODE]]$maxpops)`--`r max(structure.params.LST[[MODE]]$maxpops)`; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference). We used the variant of Evanno's K method used in \texttt{Structure Harvester} [XXX] to identify the most suitable number of populations separately for each species. We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the _G_ static; `r structure.params.LST[[MODE]]$repeats` iterations). After the \texttt{ClumPP} analysis, if most of the individuals for a given species (> 50 %) had uncertain population memberships (< 75 % probability for most likely population) then we treated this species as containing only a single population. We then ran $\texttt{Bayescan}$ for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456]. Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

>>>>

---
title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: false
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
      after_body: postamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
---

@@

## Methods
### Study system
We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

@@

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation. We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

@@

The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456]. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive. To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] using the number of populations for each species previously determined by [@r478; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference]. We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations). We then ran $\texttt{Bayescan}$ [@r456] for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations]. Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

### 2016-06-16 code/rmarkdown/figures.Rmd
+ update article with population changes + change endnote_lib.bib to references.bib

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

### 2016-06-16 code/rmarkdown/supporting_information.Rmd
+ update article with population changes + change endnote_lib.bib to references.bib

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex3.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex3.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
